---
title: "About"
output: 
  html_document: 
    theme: journal
    code_folding: hide
---

Asthma is one of the most common chronic conditions in children across the United States, currently affecting more than [6 million children](https://www.aafa.org/asthma-facts/). Ranging from mild to severe, asthma affects children's ability to breathe, exercise, and sleep. New York City (NYC) hosts a much higher prevalence of childhood asthma [(13%)](https://www1.nyc.gov/assets/doh/downloads/pdf/epi/databrief90.pdf) than the rest of the United States [(7.5%)](https://www.cdc.gov/nchs/fastats/asthma.htm) - with children of color and low-income neighborhoods affected disproportionately: 


```{r, echo = FALSE, include = FALSE}

library(tidyverse)
library(plotly)
library(tmap)
library(sf)
library(maptools)
library(leaflet)
library(RColorBrewer)

# reading in dataset
tidy_asthma = 
  read_csv("./final_data/tidy_asthma.csv") 

# reading in shapefiles
uhf_map = 
  st_read("./final_data/shapefiles/")
str(uhf_map)

# modifying vector types for mapping
tidy_asthma_sf = merge(uhf_map, tidy_asthma, by.x = "UHFCODE", by.y = "geo_join_id") %>% 
  janitor::clean_names() %>% 
  mutate(uhfcode = as.character(uhfcode),
         time_period = as.character(time_period),
         homes_with_cockroaches = as.numeric(homes_with_cockroaches))

#changing geometry to latitude and longitude
tidy_asthma_sf_ll = st_transform(tidy_asthma_sf, "+proj=longlat +datum=WGS84")

# natural breaks for cockroaches
cut(tidy_asthma$homes_with_cockroaches, 6) 
roach_bins = c(0, 5, 15, 25, 35, 40, 50)

#creating color palette for cockroaches
roach_pal = colorBin("Reds", bins = roach_bins)

# natural breaks for asthma
cut(tidy_asthma$public_school_children_5_14_yrs_old_with_asthma, 6) 
asthma_bins = c(0, 35, 55, 75, 95, 115, 150)

#color palette for asthma
asthma_pal = colorBin("Reds", bins = asthma_bins)

```

```{r, eval = TRUE}

asthma_map = 
tidy_asthma_sf_ll %>%
  filter(time_period == 2014) %>% 
  mutate(asthma_label = str_c("Children 5-14 with Asthma: ", public_school_children_5_14_yrs_old_with_asthma, "\n, Neighborhood: ", uhf_neigh)) %>% 
    leaflet() %>%
    addProviderTiles("CartoDB.Positron") %>%
    addPolygons(stroke = TRUE,
                weight = 2, 
                opacity = 1,
                color = ~asthma_pal(public_school_children_5_14_yrs_old_with_asthma),
                fillOpacity = 0.8,
                smoothFactor = 0.2,
                highlight = highlightOptions(weight = 3,
                                             color = "white",
                                             bringToFront = TRUE),
                label = ~asthma_label) %>% 
    addLegend("topleft",
              pal = asthma_pal,
              values = ~public_school_children_5_14_yrs_old_with_asthma,
              title = "Children 5-14 with Asthma (rate per 1000)",
              opacity = 1)
asthma_map
```

Aside from creating a financial [burden](https://www.ajmc.com/newsroom/cdc-study-puts-economic-burden-of-asthma-at-more-than-80-billion-per-year) for the US economy and families, asthma leads to student absenteeism, further perpetuating educational gaps and socioeconomic disparities. It has been estimated that [14 - 18%](https://www.cdc.gov/pcd/issues/2019/19_0074.htm) of missed school days are explained by asthma episodes, making asthma one of the top causes of absenteeism. 

The toll of asthma on children has drawn attention and resources towards mitigation. While root causes of asthma are still [unknown](https://www1.nyc.gov/site/doh/health/health-topics/asthma.page), common triggers - particularly those associated with allergens - have been isolated. Household pests, including cockroaches and mice, are one of the most [common indoor triggers](https://www.cdc.gov/asthma/triggers_indoor.html). Among inner city children with asthma, exposure to cockroach allergens has a [greater effect on asthma morbidity](https://rdcu.be/bX3KA) than dust mite or pet allergies. In fact, in the [United States](https://www-ncbi-nlm-nih-gov.ezproxy.cul.columbia.edu/pmc/articles/PMC3261483/), the rate of hospitalization for asthma among children living in inner-cities homes with cockroach allergens is 3.4 times higher than non-sensitized/non-exposed children.

The analyses presented here explore the association between childhood asthma and another highly prevalent NYC phenomenon - cockroaches. In 2017, 15.6% of households [reported](http://www.sternenvironmental.com/blog/2017/04/18/nyc-breaks-the-top-ten-list-for-worst-rat-and-cockroach-problems/) having pest infestations. As temperatures rise, cockroach populations are [expected](https://gothamist.com/arts-entertainment/its-summer-in-nyc-and-the-cockroaches-have-taken-flight) to increase, posing additional public health challenges, particularly among those who have asthma or respiratory reactions to cockroaches. The prevalence of homes with cockroaches, measured as percent of households in an area that reported seeing at least one cockroach daily over the last month divided by number households in the area, is shown below. 


```{r, eval = TRUE}

roach_map = 
tidy_asthma_sf_ll %>%
  filter(time_period == 2014) %>% 
  mutate(roach_label = str_c("Homes with Cockroaches: ", homes_with_cockroaches, "\n, Neighborhood: ", uhf_neigh)) %>% 
    leaflet() %>%
    addProviderTiles("CartoDB.Positron") %>%
    addPolygons(stroke = TRUE,
                weight = 2, 
                opacity = 1,
                color = ~roach_pal(homes_with_cockroaches),
                fillOpacity = 0.8,
                smoothFactor = 0.2,
                highlight = highlightOptions(weight = 3,
                                             color = "white",
                                             bringToFront = TRUE),
                label = ~roach_label) %>% 
    addLegend("topleft",
              pal = roach_pal,
              values = ~homes_with_cockroaches,
              title = "Homes with Cockroaches",
              opacity = 1, 
              labFormat = function(type, cuts, p) {
                n = length(cuts)
                paste0(round(cuts, 1)[-n], " &ndash; ", round(cuts, 1)[-1], "%")
                })
roach_map
```




