---
title: "alice data review"
author: "Alice Tivarovsky"
date: "11/15/2019"
output: html_document
---

## Setup
```{r setup}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## Data import and tidying

Loading in and tidying data: 
```{r}

asthma_data = read_csv(file = "./final_data/asthma_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(data_value != "NULL") %>% 
  rename(year = year_description) %>% 
  mutate(
    data_value = as.numeric(data_value),
    year = as.factor(year)
    ) %>% 
  drop_na(year) %>% 
  filter(str_detect(name, c("[Aa]sthma", "[Cc]ockroaches"))) %>% 
  select(year, geo_entity_id, name, data_value) %>% 
  pivot_wider(
    names_from = name,
    values_from = data_value
  ) %>% 
  view()
 
```

Replacing NA values with averages across years using a function: 

```{r}

asthma_replace = function(x){
  
  y = mean(x, na.rm = TRUE)
  x = replace(x, is.na(x), y)
  
}

```

```{r}

asthma_data %>% 
  group_by(geo_entity_id, year) %>% 
  summarize(n = mean("Adults with Asthma in the Past 12 Months", na.rm = TRUE))

```


We removed entries that included non-numerical year values, missing values for data_value, and any indicator not containing the words "asthma", "cockroaches", or "poverty". The resulting dataset contains 2,212 entries across 9 variables. Variables are defined as follows: 

* `indicator_data_id`: unique data entry id
* `name`: 


## Trying with all available data 

```{r}

all_data = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Adults with Asthma in the Past 12 Months",
                     "Asthma Emergency Department Visits (Adults)", 
                     "Asthma Emergency Department Visits (Children 0 to 4 Yrs Old)", 
                     "Asthma Emergency Department Visits (Children 5 to 17 Yrs Old)", 
                     "Asthma Hospitalizations (Adults)", 
                     "Asthma Hospitalizations (Children 0 to 4 Yrs Old)", 
                     "Asthma Hospitalizations (Children 5 to 17 Yrs Old)", 
                     "Children under 5 years old in Poverty", 
                     "Homes with Cockroaches", 
                     "Neighborhood Poverty", 
                     "Neighborhood Poverty (ACS)", 
                     "O3-Attributable Asthma Emergency Department Visits", 
                     "O3-Attributable Asthma Hospitalizations ",
                     "PM2.5-Attributable Asthma Emergency Department Visits ", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma"),
        geo_type_name == "UHF42", 
        time_period %in% c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017)
         ) %>% 
  distinct()
  

```

Other way: 
```{r}
all_data_2 = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(name = as.character(name)) %>% 
  filter(str_detect(name, c("[Aa]sthma", "[Cc]ockroaches", "[Pp]overty")),
         str_detect(name, "Adults Reporting Cockroaches in the Home", negate = TRUE),
         str_detect(name, "NTA", negate = TRUE),
         geo_type_name == 'UHF42'
         ) %>% 
  distinct()
```


```{r}
 pull(distinct(all_data, measure))
```

Exploratory counts to verify filter steps: 

```{r}
all_data %>%  
  group_by(name, time_period, measure, geo_join_id) %>% 
  summarize(data_value = sum(data_value)) %>% 
  view()
```



Tidying: 

```{r}
all_data %>%
pivot_wider(
    names_from = name,
    values_from = data_value
  ) %>% 
view()
```


## Tidy for school children

```{r}
all_data = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Homes with Cockroaches", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma"),
        geo_type_name == "UHF42", 
        measure %in% c("Percent", "Rate")
        #time_period %in% c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017)
         ) %>% 
  pivot_wider(
    names_from = name, 
    values_from = data_value
    ) %>% 
  view()
```

```{r}
 pull(distinct(all_data, name))
```

For Homes with Cockroaches, Percent = estimated # of households in an area that reported seeing at least one cockroach daily over the last month divided by # households in the area

For Poverty, Percent = estimated # of ppl for whom poverty status is determined whose annual income falls below 100% of FPL divided by # of ppl for whom poverty status is determined. 

For public school children with asthma rates: # of public school children with diagnosis of asthma during current or previous school year divided by 5-14 year old population expresssed per 1000. 



```{r}
all_data  %>% 
  select(geo_join_id, Poverty, time_period) %>% 
  ggplot(aes(x = geo_join_id, y = Poverty, color = time_period)) +
  geom_line(aes(group = time_period))


```

### Smooshing: 

```{r}
all_data_3 = 
  all_data %>% 
  mutate(
    time_period = recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011")) %>%
  filter(time_period %in% c("2011", "2012", "2013", "2014")) %>% 
  janitor::clean_names() %>% 
  select(geo_join_id, time_period, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma) %>%
  group_by(geo_join_id, time_period) %>% 
  summarize()

all_data_4 = 
  all_data %>% 
  mutate(
    time_period = recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011")) %>%
  filter(time_period %in% c("2011", "2012", "2013", "2014")) %>% 
  janitor::clean_names() %>% 
  select(geo_join_id, time_period, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma)
  
all_data_5 =
 inner_join(all_data_2, all_data_3, by = "geo_join_id", "time_period")
```


Trying to merge with shapefiles 
```{r}

#shape = st_read("./final_data/shapefiles/", quiet = TRUE)
#merged <- merge(all_data, shape, by.x = "geo_join_id", by.y = "UHFCODE")


#merged %>% filter(time_period == "2012") %>% view()

#plot(st_geometry(shape))
```



```{r}
all_data_2 %>% 
  group_by(time_period) %>% 
  summarize(n = n())

```


