---
title: "Kelly"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggridges)
library(patchwork)
library(plotly)
library(ggplot2)
library(gganimate)
library(modelr)
library(mgcv)

```

```{r, error = TRUE}
all_available_data=
  read_csv(file="./final_data/all_available_data.csv")

Asthma_data=
  read_csv(file="./final_data/asthma_data.csv")

mold_data=
  read_csv(file="./final_data/mold_home.csv")
```

### Kelly vision: X : borough name , Y: Cockroach Rate, by different boroughs 
#### Trying to include Borough's in here--using Alice's Method to include Borough name

#### Vision did not work because did not really have enough time lapse for others. 
```{r, error = TRUE}
overview_plot=
all_available_data %>% 
  filter(Name= "Public_School_Children_ ())
```

```{r, error = TRUE}
pp = tidy_asthma %>%
  drop_na(homes_with_cockroaches) %>% 
  mutate(
    geo_join_id=as.numeric(geo_join_id) ) %>% 
  ggplot(aes(x = geo_join_id, y = homes_with_cockroaches)) + 
  geom_point(aes(colour = time_period), alpha = 0.5) +
  labs(title = "Prevalence of Cockroaches in Different Neighborhoods",
     x = 'Geo_ID', 
     y = 'Rate of Homes with Cockroaches(%)'
)
pp

# gganimate parts
pp + transition_time(time_period) + labs(title="Year:{frame_time}")

animate(pp, fps = 2, height = 600, width = 1000)
```

## using lorraine's code as a guide
```{r, error = TRUE}
test_prev=
  tidy_asthma %>% 
  drop_na(homes_with_cockroaches) %>% 
  mutate(
    geo_join_id=as.numeric(geo_join_id),
    time_period=as.factor(time_period),
    time_period=fct_relevel(time_period, "2011"="2011", "2012"="2012", "2013" = "2013", "2014" = "2014")
  ) %>% 
  arrange(geo_join_id) %>% 
  plot_ly(
    x=~geo_join_id, 
    y=~homes_with_cockroaches,
    color=~geo_join_id,
    frame=~ time_period,
    hoverinfo= "text",
    type="scatter",
    mode="markers"
  )

test_prev
  
```

## attempting to look at hospitalizations and Using Alice's Code
```{r, error = TRUE}
all_data = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Homes with Cockroaches", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma",
                     "Asthma Hospitalizations (Children 5 to 17 Yrs Old)"),
        geo_type_name == "UHF42", 
        measure %in% c("Percent", "Rate"),
        geo_place_name
         ) %>% 
  pivot_wider(
    names_from = name, 
    values_from = data_value
    ) %>% 
  view()

all_data=
  all_data %>% 
  mutate(
    time_period = recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011")) %>%
  filter(time_period %in% c("2011", "2012", "2013", "2014")) %>% 
  janitor::clean_names() %>% 
  select(geo_join_id, time_period, measure, indicator_id, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma,asthma_hospitalizations_children_5_to_17_yrs_old, geo_place_name )
````
```{r}
## merging using Alice's method
df1 = 
  all_data  %>% 
  select(geo_join_id, time_period, measure, indicator_id, homes_with_cockroaches) %>% 
  drop_na()


df2 = 
all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, poverty) %>% 
  drop_na()
df2

df3 = 
  all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, public_school_children_5_14_yrs_old_with_asthma) %>% 
  drop_na()
df3

df4 = 
  all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, public_school_children_5_14_yrs_old_with_persistent_asthma) %>% 
  drop_na()
df4
df5 = 
  all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  drop_na()
df5


## Joining
tidy_asthma_hospitalization = 
  full_join(df1, df2, by = c("geo_join_id", "time_period")) %>% 
  full_join(df3, by = c("geo_join_id", "time_period")) %>% 
  full_join(df4, by = c("geo_join_id", "time_period")) %>% 
  full_join(df5, by= c("geo_join_id", "time_period")) %>% 
  select(geo_join_id, time_period, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma, asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  view()
```

```{r, error = TRUE}
## see changes with Hospitalzation by time

test_prev=
  tidy_asthma_hospitalization %>% 
  drop_na(asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  mutate(
    geo_join_id=as.character(geo_join_id),
   time_period = as.numeric(levels(time_period)) [time_period]
  ) %>% 
  arrange(geo_join_id) %>% 
  plot_ly(
    x=~geo_join_id, 
    y=~asthma_hospitalizations_children_5_to_17_yrs_old,
    color=~geo_join_id,
    frame=~ time_period,
    hoverinfo= "text",
    type="scatter",
    mode="markers"
  )

test_prev
```
<<<<<<< HEAD


## regression model:
```{r, error = TRUE}
#set seed for reproducibility
set.seed(10)
#fitting the full model


tidy_asthma = 
  read_csv(file = "./final_data/tidy_asthma.csv")
test=
  tidy_asthma %>% 
  mutate(
    public_school_children_5_14_yrs_old_with_asthma=as.numeric(public_school_children_5_14_yrs_old_with_asthma),
    homes_with_cockroaches=as.numeric(homes_with_cockroaches),
    poverty=as.numeric(poverty),
    geo_join_id=as.numeric(geo_join_id),
    geo_place_name=as.factor(geo_place_name)
  )



## Model 1
model1_fit = lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches + poverty + geo_join_id, data=test) %>% broom::tidy()

model1_fit

## Model 2
model2_fit = lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * geo_join_id, data=test) %>% broom::tidy()

###### Time_Period doesn't work
##Model 3
model3_fit = lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * geo_join_id*time_period, data=test) %>% broom::tidy()

##Model4
model4_fit = lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches*poverty*time_period, data=test) %>% broom::tidy()

## Model 5
model5_fit = lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * geo_join_id * asthma_hospitalizations_children_5_to_17_yrs_old, data=test) %>% broom::tidy()

# Looking at residuals 
residual_1=
  modelr::add_residuals(tidy_asthma, model1_fit) %>% 
  ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma, y = resid)) + geom_point()

residual_2=
  modelr::add_residuals(tidy_asthma, model2_fit) %>% 
  ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma, y = resid)) + geom_point()

residual_3=
  modelr::add_residuals(tidy_asthma, model3_fit) %>% 
  ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma, y = resid)) + geom_point()

residual_4=
  modelr::add_residuals(tidy_asthma, model4_fit) %>% 
  ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma, y = resid)) + geom_point()

residual_5=
  modelr::add_residuals(tidy_asthma, model5_fit) %>% 
  ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma, y = resid)) + geom_point()


ggplot_data = 
tidy_asthma %>%
  filter(time_period == c("2011", "2014"))
  
linear_model = 
  lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * time_period * geo_join_id *asthma_hospitalizations_children_5_to_17_yrs_old, data = ggplot_data) %>%
  broom::tidy() 

plot_tidy = 
ggplot(ggplot_data, aes(x = homes_with_cockroaches, y = public_school_children_5_14_yrs_old_with_asthma, color = geo_join_id)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_grid(~time_period) +
  labs(x = "Percent of Homes with Cockroaches",
       y = "Percent of School Children with Asthma",
       title = "Effect of Homes with Cockroaches on Childhood Asthma adjusting for Neighborhood") +
  viridis::scale_color_viridis(aes(color = geo_join_id), discrete = FALSE)

plot_tidy

```

## Cross Validation
```{r, error = TRUE}
cv_df=
  crossv_mc(tidy_asthma, 100) %>% 
  mutate(
    train=map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_df=
  cv_df %>% 
  mutate(fit = map(train, ~lm(public_school_children_5_14_yrs_old_with_asthma ~ ., data=.x)),
        model1_lm= map(train, ~lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches + poverty + geo_join_id, data=.x)),
        model2_lm=map(train, ~lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * geo_join_id, data=.x)),
        model5_lm=map(train, ~lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * geo_join_id * asthma_hospitalizations_children_5_to_17_yrs_old, data=.x))) %>% 
  mutate(rmse_fit = map2_dbl(fit, test, ~rmse(model=.x, data=.y)),
         rmse_model1= map2_dbl(model1_lm, test, ~rmse(model=.x, data=.y)),
         rmse_model2= map2_dbl(model2_lm, test, ~rmse(model=.x, data=.y)),
         rmse_model5=map2_dbl(model5_lm, test, ~rmse(model=.x, data=.y)))
```

```

=======
>>>>>>> 47a2c4709bdc3516aa253444ade91a33d5b9a8a9
