---
title: "lorraine"

output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(viridis)
library(sp)
library(rvest)
library(tigris)
library(maptools)
library(ggmap)
library(broom)
library(httr)
library(rgdal)
library(readxl)
library(patchwork)
library(plotly)
library(gganimate)
library(magick)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Most up to date tidy dataset

```{r tidying data, message=FALSE, include=FALSE, error = TRUE}
all_data = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Homes with Cockroaches", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma", 
                     "Asthma Hospitalizations (Children 5 to 17 Yrs Old)"),
        geo_type_name %in% c("UHF42", "Borough"), 
        measure %in% c("Percent", "Rate")
         ) %>% 
  mutate(geo_join_id = as.character(geo_join_id),
         geo_join_id = if_else(geo_place_name == "Bensonhurst - Bay Ridge", "209", geo_join_id),
         geo_join_id = as.factor(geo_join_id)) %>% 
  pivot_wider(
    names_from = name, 
    values_from = data_value
    ) %>%
  mutate(time_period = recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011")) %>%
  filter(time_period %in% c("2011", "2012", "2013", "2014")) %>% 
  janitor::clean_names() %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma, asthma_hospitalizations_children_5_to_17_yrs_old)

df1 = all_data  %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, homes_with_cockroaches) %>% 
  drop_na()

df2 = all_data %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, poverty) %>% 
  drop_na()

df3 = all_data %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, public_school_children_5_14_yrs_old_with_asthma) %>% 
  drop_na()

df4 = all_data %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, public_school_children_5_14_yrs_old_with_persistent_asthma) %>% 
  drop_na()

df5 = all_data %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  drop_na()

## Joining
tidy_asthma = 
  full_join(df1, df2, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df3, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df4, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df5, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma, asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  mutate(time_period = as.numeric(levels(time_period)) [time_period],
         geo_join_id = as.character(as.numeric(levels(geo_join_id)) [geo_join_id]),
         geo_place_name = as.character(geo_place_name)) %>%
  view()

write.csv(tidy_asthma,"./final_data/tidy_asthma.csv", row.names = FALSE)
```

PLOTS FOR PUBLIC SCHOOL CHILDREN VARIABLES

SCATTERPLOTS

```{r}
plot_uhf_asthma =
  tidy_asthma %>%
  filter(geo_type_name == "UHF42") %>%
  arrange(geo_join_id) %>%
    plot_ly(
    x = ~geo_join_id, 
    y = ~public_school_children_5_14_yrs_old_with_asthma, 
    color = ~geo_place_name,
    text = ~geo_place_name,
    frame = ~time_period,
    hoverinfo = "text",
    type = 'scatter',
    mode = 'markers'
  )
  
plot_uhf_asthma
```

```{r}
plot_uhf_persistent =
  tidy_asthma %>%
  filter(geo_type_name == "UHF42") %>%
  arrange(geo_join_id) %>%
  plot_ly(
    x = ~geo_join_id, 
    y = ~public_school_children_5_14_yrs_old_with_persistent_asthma, 
    color = ~geo_place_name,
    text = ~geo_place_name,
    frame = ~time_period,
    hoverinfo = "text",
    type = 'scatter',
    mode = 'markers'
  )

plot_uhf_persistent
```

```{r}
subplot(plot_uhf_asthma, plot_uhf_persistent)
```

testing different types of plots for asthma variables

BAR CHARTS

```{r}
bar_uhf_asthma =
  tidy_asthma %>%
  filter(geo_type_name == "UHF42") %>%
  arrange(geo_join_id) %>%
    plot_ly(
    x = ~geo_join_id, 
    y = ~public_school_children_5_14_yrs_old_with_asthma, 
    color = ~geo_place_name,
    text = ~geo_place_name,
    frame = ~time_period,
    hoverinfo = "text",
    type = 'bar'
  )
  
bar_uhf_asthma
```

```{r}
bar_uhf_persistent =
  tidy_asthma %>%
  filter(geo_type_name == "UHF42") %>%
  arrange(geo_join_id) %>%
  plot_ly(
    x = ~geo_join_id, 
    y = ~public_school_children_5_14_yrs_old_with_persistent_asthma, 
    color = ~geo_place_name,
    text = ~geo_place_name,
    frame = ~time_period,
    hovertemplate = paste(
      "<b>%{text}</b><br>",
      "<i>Rate</i>: %{y:.2f}"),
    type = 'bar'
  )

bar_uhf_persistent
```

```{r}
plot_borough_asthma =
  tidy_asthma %>%
  filter(geo_type_name == "Borough") %>%
  arrange(geo_join_id) %>%
  plot_ly(
    x = ~geo_place_name, 
    y = ~public_school_children_5_14_yrs_old_with_asthma, 
    color = ~geo_place_name,
    text = ~geo_place_name,
    frame = ~time_period,
    hovertemplate = paste(
      "<b>%{x}</b><br>",
      "<i>Rate</i>: %{y:.2f}"),
    type = 'bar'
  )

plot_borough_asthma
```

```{r}
plot_borough_persistent =
  tidy_asthma %>%
  filter(geo_type_name == "Borough") %>%
  arrange(geo_join_id) %>%
  plot_ly(
    x = ~geo_place_name, 
    y = ~public_school_children_5_14_yrs_old_with_persistent_asthma, 
    color = ~geo_place_name,
    text = ~geo_place_name,
    frame = ~time_period,
    hovertemplate = paste(
      "<b>%{x}</b><br>",
      "<i>Rate</i>: %{y:.2f}"),
    type = 'bar'
  )

plot_borough_persistent
```

```{r}
plot_borough_cockroach =
  tidy_asthma %>%
  filter(geo_type_name == "Borough") %>%
  arrange(geo_join_id) %>%
  plot_ly(
    x = ~geo_place_name, 
    y = ~homes_with_cockroaches, 
    color = ~geo_place_name,
    text = ~geo_place_name,
    frame = ~time_period,
    hovertemplate = paste(
      "<b>%{x}</b><br>",
      "<i>Percent</i>: %{y:.2f}"),
    type = 'bar'
  )

plot_borough_cockroach
```

test_2 = 
  tidy_asthma %>%
  ggplot(aes(x = geo_join_id, y = public_school_children_5_14_yrs_old_with_asthma, frame = time_period)) +
geom_point(aes(colour = geo_join_id, alpha = 0.5)) + 
labs(title = "Test",
x = 'UHF Code', 
y = 'Public School Children (5-14 Years Old) with Asthma (%)',
colour = 'Year') +
theme(plot.title = element_text(size = 40, face = "bold"),
axis.text = element_text(size = 18),
axis.title = element_text(size = 18,face = "bold")) + 
theme(legend.text = element_text(size = 16), legend.title = element_text(size = 18,face = "bold") ) +
transition_states(time_period, transition_length = 1, state_length = 3, wrap = TRUE) +
  enter_fade() +
  exit_fade()

animate(test_2, fps = 2, height = 600, width = 1000)


    