---
title: "lorraine"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(viridis)
library(sp)
library(rvest)
library(tigris)
library(maptools)
library(ggmap)
library(broom)
library(httr)
library(rgdal)
library(readxl)
library(patchwork)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Reading in the csv file. 

```{r}
asthma_data = read.csv(file = "./final_data/asthma_data.csv") %>%
  janitor::clean_names() %>%
  mutate(
    data_value = as.numeric(data_value), 
    year = as.character(year_description)) %>%
  select(year, geo_entity_id, name, data_value) %>%
  pivot_wider(names_from = name,
              values_from = data_value) %>%
  janitor::clean_names() %>%
  view()
```

Loading and tidying the new dataset with all available data. THis generates 4536 observations. 

```{r}
all_data_1 = 
  read.csv(file = "./final_data/all_available_data.csv") %>%
  janitor::clean_names() %>%
  rename(year = time_period) %>%
  mutate(
    data_value = as.numeric(data_value), 
    geo_join_id = as.numeric(geo_join_id),
    year = as.character(year), 
    name = as.character(name)) %>%
  filter(data_value != "NULL",
         str_detect(name, c("[A]sthma", "[C]ockroach", "[P]overty")),
         !str_detect(name, c("[N]TA")),
         name != "Children 0-13 Years Old Ever Diagnosed with Asthma",
         name != "Youth Asthma in the Past 12 Months",
         name != "Adults Reporting Cockroaches in the Home",
         geo_type_name == "UHF42") %>%
  group_by(name) %>%
  janitor::clean_names() %>%
  view()
```

Tidying up the data using factors instead of strings for the `name` variable. 

```{r}
all_data = 
  read.csv(file = "./final_data/all_available_data.csv") %>%
  janitor::clean_names() %>%
  mutate(data_value = as.numeric(data_value)) %>%
  filter(name %in% c("Adults with Asthma in the Past 12 Months",
                     "Asthma Emergency Department Visits (Adults)", 
                     "Asthma Emergency Department Visits (Children 0 to 4 Yrs Old)", 
                     "Asthma Emergency Department Visits (Children 5 to 17 Yrs Old)", 
                     "Asthma Hospitalizations (Adults)", 
                     "Asthma Hospitalizations (Children 0 to 4 Yrs Old)", 
                     "Asthma Hospitalizations (Children 5 to 17 Yrs Old)", 
                     "Children under 5 years old in Poverty", 
                     "Homes with Cockroaches", 
                     "Neighborhood Poverty", 
                     "Neighborhood Poverty (ACS)", 
                     "O3-Attributable Asthma Emergency Department Visits", 
                     "O3-Attributable Asthma Hospitalizations ",
                     "PM2.5-Attributable Asthma Emergency Department Visits ", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma"),
         geo_type_name == "UHF42")
```

Filtering for variables of interest (1344 observations) 

```{r}
all_data %>%
  filter(name %in% c("Homes with Cockroaches",
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma")) %>%
  mutate(row = row_number()) %>%
  select(row, name, measure, geo_join_id, time_period, data_value) %>%
  pivot_wider(names_from = "name",
              values_from = "data_value") %>%
  select(-row) %>%
  view()
```

```{r}
all_data %>%
  filter(name %in% c("Homes with Cockroaches",
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma")) %>%
  mutate(row = row_number()) %>%
  select(row, name, measure, geo_join_id, time_period, data_value) %>%
  pivot_wider(names_from = "time_period",
              values_from = "data_value") %>%
  select(-row) %>%
  view()
```

Attempting to create function to fill in NAs - Example: we want to use the 2008 data for cockroaches to fill in the missing values for any split-years that contain 2008 in the range. 



