---
title: "lorraine"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(viridis)
library(sp)
library(rvest)
library(tigris)
library(maptools)
library(ggmap)
library(broom)
library(httr)
library(rgdal)
library(readxl)
library(patchwork)
library(plotly)
library(gganimate)
library(magick)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Reading in the csv file. 

```{r}
asthma_data = read.csv(file = "./final_data/asthma_data.csv") %>%
  janitor::clean_names() %>%
  mutate(
    data_value = as.numeric(data_value), 
    year = as.character(year_description)) %>%
  select(year, geo_entity_id, name, data_value) %>%
  pivot_wider(names_from = name,
              values_from = data_value) %>%
  janitor::clean_names() %>%
  view()
```

Loading and tidying the new dataset with all available data. THis generates 4536 observations. 

```{r}
all_data_1 = 
  read.csv(file = "./final_data/all_available_data.csv") %>%
  janitor::clean_names() %>%
  rename(year = time_period) %>%
  mutate(
    data_value = as.numeric(data_value), 
    geo_join_id = as.numeric(geo_join_id),
    year = as.character(year), 
    name = as.character(name)) %>%
  filter(data_value != "NULL",
         str_detect(name, c("[A]sthma", "[C]ockroach", "[P]overty")),
         !str_detect(name, c("[N]TA")),
         name != "Children 0-13 Years Old Ever Diagnosed with Asthma",
         name != "Youth Asthma in the Past 12 Months",
         name != "Adults Reporting Cockroaches in the Home",
         geo_type_name == "UHF42") %>%
  group_by(name) %>%
  janitor::clean_names() %>%
  view()
```

Tidying up the data using factors instead of strings for the `name` variable. 

```{r}
all_data = 
  read.csv(file = "./final_data/all_available_data.csv") %>%
  janitor::clean_names() %>%
  mutate(data_value = as.numeric(data_value),
         time_period = recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011"),
         geo_join_id = as.numeric(geo_join_id)) %>%
  filter(name %in% c("Homes with Cockroaches",
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma"),
                     measure == "Percent" | measure == "Rate",
         geo_type_name == "UHF42",
         time_period %in% c("2011", "2012", "2013", "2014")) %>%
  mutate(row = row_number(),
         time_period = as.numeric(time_period)) %>%
  pivot_wider(names_from = "name",
              values_from = "data_value") %>%
  janitor::clean_names() %>%
  select(geo_join_id, time_period, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma) %>%
  view()
```

Filtering for variables of interest (1344 observations) 

```{r}
test =
  all_data %>%
  group_by(time_period, geo_join_id) %>%
  summarise_each(funs(sum(., na.rm = TRUE))) %>% 
  view()
```

```{r}
all_data %>%
  select(geo_join_id, poverty, time_period) %>%
  ggplot(aes(x = geo_join_id, y = poverty, color = time_period))  +
  geom_line(aes(group = time_period))
```


USE THIS: 

```{r}
all_data = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Adults with Asthma in the Past 12 Months",
                     "Asthma Emergency Department Visits (Adults)", 
                     "Asthma Emergency Department Visits (Children 0 to 4 Yrs Old)", 
                     "Asthma Emergency Department Visits (Children 5 to 17 Yrs Old)", 
                     "Asthma Hospitalizations (Adults)", 
                     "Asthma Hospitalizations (Children 0 to 4 Yrs Old)", 
                     "Asthma Hospitalizations (Children 5 to 17 Yrs Old)", 
                     "Children under 5 years old in Poverty", 
                     "Homes with Cockroaches", 
                     "Neighborhood Poverty", 
                     "Neighborhood Poverty (ACS)", 
                     "O3-Attributable Asthma Emergency Department Visits", 
                     "O3-Attributable Asthma Hospitalizations ",
                     "PM2.5-Attributable Asthma Emergency Department Visits ", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma"),
        geo_type_name == "UHF42", 
        time_period %in% c(2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017)
         ) %>% 
  distinct()
```

Other way: 

```{r}
all_data_2 = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(name = as.character(name)) %>% 
  filter(str_detect(name, c("[Aa]sthma", "[Cc]ockroaches", "[Pp]overty")),
         str_detect(name, "Adults Reporting Cockroaches in the Home", negate = TRUE),
         str_detect(name, "NTA", negate = TRUE),
         geo_type_name == 'UHF42'
         ) %>% 
  distinct()
```


```{r}
 pull(distinct(all_data, measure))
```

Exploratory counts to verify filter steps: 

```{r}
all_data %>%  
  group_by(name, time_period, measure, geo_join_id) %>% 
  summarize(data_value = sum(data_value)) %>% 
  view()
```

Tidying: 

```{r}
all_data %>%
pivot_wider(
    names_from = name,
    values_from = data_value
  ) %>% 
view()
```

## Tidy for school children only - USE THIS 

```{r}
all_data = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Homes with Cockroaches", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma"),
        geo_type_name == "UHF42", 
        measure %in% c("Percent", "Rate")
         ) %>% 
  pivot_wider(
    names_from = name, 
    values_from = data_value
    ) %>% 
  view()
```

For Homes with Cockroaches, Percent = estimated # of households in an area that reported seeing at least one cockroach daily over the last month divided by # households in the area

For Poverty, Percent = estimated # of ppl for whom poverty status is determined whose annual income falls below 100% of FPL divided by # of ppl for whom poverty status is determined. 

For public school children with asthma rates: # of public school children with diagnosis of asthma during current or previous school year divided by 5-14 year old population expresssed per 1000. 


Exploring differences in poverty by time: 

```{r}
all_data  %>% 
  select(geo_join_id, Poverty, time_period) %>% 
  ggplot(aes(x = geo_join_id, y = Poverty, color = time_period)) +
  geom_line(aes(group = time_period))
```
Conclude not much difference. Can recode years


### Recoding years: 

```{r}
all_data = 
  all_data %>% 
  mutate(
    time_period = recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011")) %>%
  filter(time_period %in% c("2011", "2012", "2013", "2014")) %>% 
  janitor::clean_names() %>% 
  select(geo_join_id, time_period, measure, indicator_id, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma)
```

Trying Jiayi's suggestion to get rid of stacking: 

```{r}
df1 = 
  all_data  %>% 
  select(geo_join_id, time_period, measure, indicator_id, homes_with_cockroaches) %>% 
  drop_na()


df2 = 
all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, poverty) %>% 
  drop_na()
df2

df3 = 
  all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, public_school_children_5_14_yrs_old_with_asthma) %>% 
  drop_na()
df3

df4 = 
  all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, public_school_children_5_14_yrs_old_with_persistent_asthma) %>% 
  drop_na()
df4

## Joining
tidy_asthma = 
  full_join(df1, df2, by = c("geo_join_id", "time_period")) %>% 
  full_join(df3, by = c("geo_join_id", "time_period")) %>% 
  full_join(df4, by = c("geo_join_id", "time_period")) %>% 
  select(geo_join_id, time_period, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma) %>%
  mutate(time_period = as.numeric(levels(time_period)) [time_period]) %>%
  view()
  
```

PLOTS FOR PUBLIC SCHOOL CHILDREN VARIABLES

```{r}
p1 = 
  tidy_asthma %>%
  filter(time_period == "2011") %>%
  ggplot(aes(x = geo_join_id, y = public_school_children_5_14_yrs_old_with_asthma, color = geo_join_id)) + 
  geom_point()

p2 = 
  tidy_asthma %>%
  filter(time_period == "2012") %>%
  ggplot(aes(x = geo_join_id, y = public_school_children_5_14_yrs_old_with_asthma, color = geo_join_id)) + 
  geom_point()

p3 = 
  p1 = 
  tidy_asthma %>%
  filter(time_period == "2013") %>%
  ggplot(aes(x = geo_join_id, y = public_school_children_5_14_yrs_old_with_asthma, color = geo_join_id)) + 
  geom_point()

p4 = 
  tidy_asthma %>%
  filter(time_period == "2014") %>%
  ggplot(aes(x = geo_join_id, y = public_school_children_5_14_yrs_old_with_asthma, color = geo_join_id)) + 
  geom_point()
```

```{r}
test_1 =
  tidy_asthma %>%
  plot_ly(
    x = ~geo_join_id, 
    y = ~public_school_children_5_14_yrs_old_with_asthma, 
    color = ~geo_join_id,
    text = ~geo_join_id,
    frame = ~time_period,
    hoverinfo = "text",
    type = 'scatter',
    mode = 'markers'
  )
```

```{r}
test_2 = 
  tidy_asthma %>%
  ggplot(aes(x = geo_join_id, y = public_school_children_5_14_yrs_old_with_asthma, frame = time_period)) +
geom_point(aes(colour = geo_join_id, alpha = 0.5)) + 
labs(title = "Test",
x = 'UHF Code', 
y = 'Public School Children (5-14 Years Old) with Asthma (%)',
colour = 'Year') +
theme(plot.title = element_text(size = 40, face = "bold"),
axis.text = element_text(size = 18),
axis.title = element_text(size = 18,face = "bold")) + 
theme(legend.text = element_text(size = 16), legend.title = element_text(size = 18,face = "bold") ) +
transition_states(time_period, transition_length = 1, state_length = 3, wrap = TRUE) +
  enter_fade() +
  exit_fade()

animate(test_2, fps = 2, height = 600, width = 1000)
```

    