---
title: "Overview"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

## Cockroach and Asthma Distributions by Borough

```{r, echo = FALSE, include = FALSE}
library(tidyverse)
library(PerformanceAnalytics)
library(sf)
library(sp)
library(tmap)
library(patchwork)

tidy_asthma = 
  read_csv(file = "./final_data/tidy_asthma.csv")

tidy_asthma 

shape = st_read("./final_data/shapefiles/", quiet = TRUE)

tidy_asthma_sf = merge(shape, tidy_asthma, by.x = "UHFCODE", by.y = "geo_join_id") %>% 
  janitor::clean_names() %>% 
  mutate(uhfcode = as.character(uhfcode),
         time_period = as.character(time_period),
         homes_with_cockroaches = as.numeric(homes_with_cockroaches))

tidy_asthma_sf_ll = st_transform(tidy_asthma_sf, "+proj=longlat +datum=WGS84")
```

```{r, echo = FALSE}

p2 = 
  tidy_asthma_sf_ll %>%
  ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma)) + 
  geom_density() +
  facet_grid(~borough) +
  labs(
    x = "Percent of Public School Children with Asthma",
    y = "Frequency",
    title = "Distribution of Childhood Asthma by Borough")
p2
```

## Cockroach Prevalence by Borough Graph 
```{r, echo = FALSE}
p1 = 
  tidy_asthma_sf_ll %>%
  ggplot(aes(x = homes_with_cockroaches)) + 
  geom_density() +
  facet_grid(~borough) +
  labs(
    x = "Percent of Homes with Cockroaches",
    y = "Frequency",
    title = "Distribution of Cockroaches by Borough"
  ) 

p1
```


### Related work 

## Initial Questions 

## Data 

# Data

We built our dataset using the [NYC Gov Environmental & Health Data Portal](http://a816-dohbesp.nyc.gov/IndicatorPublic/BuildATable.aspx). The Environmental & Health Data Portal houses data from a variety of data sources such as the New York City Automated School Health Records, American Community Survey and New York City Housing and Vacancy Survey (HVS). Data from this portal are used to study how the environment affects health. The map was built with the United Hospital Fund (UHF) neighborhood code key from [NYC Health](http://a816-dohbesp.nyc.gov/IndicatorPublic/EPHTPDF/uhf42.pdf). The asthma CSV file we used is linked (here)[asthma_data.csv]. We selected all variables relating to asthma, cockroaches, and poverty through string detection:

  * `geo_type_name`: 
  
  * `geo_join_id`:
  
  * `geo_place_name`:
  
  * `time_period`: Year
  
  * `homes_with_cockroaches` (percent): Estimated number of households in an 
    area that reported seeing at least one cockroach daily over the last 
    month, divided by the number of households in the area.
    
  * `poverty`:
  
  * `public_school_children_5_14_yrs_old_with_asthma` (rate): Number of 
    public school children, aged 5-14 years, with a diagnosis of asthma that 
    has been reported to NYC DOHMH during the current or previous school 
    year, divided by the 5-14 year old population, per 1,000. 



* Our github repository is linked [here](https://github.com/prisrinivasan/p8105_final)

## Exploratory analysis

* The graphs and analyses can be found [here](regression.Rmd).
* We started by creating a correlation matrix to assess multicollinearity and baseline correlations
* We then assessed the crude relationship between exposure, cockroach presence in the home and the outcome, childhood asthma
* 

## Discussion

