---
title: "Overview"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Cockroach Prevalence in New York City
## Why should we care? 
In 2017, 15.6% of households [reported](http://www.sternenvironmental.com/blog/2017/04/18/nyc-breaks-the-top-ten-list-for-worst-rat-and-cockroach-problems/) having pest infestations. One of the most common pests in NYC is cockroaches. As temperatures rise, cockroach populations are [expected](https://gothamist.com/arts-entertainment/its-summer-in-nyc-and-the-cockroaches-have-taken-flight) to increase, posing additional public health challenges, particularly among those who have asthma or respiratory reactions to cockroaches.  

Because of this, we want to take a life course perspective and look at how hospitalization rates among children ages 0-4 are affected by prevalence of cockroaches in their homes.

* Insert Map Here 

## Cockroach and Asthma Distributions by Borough

```{r, echo = FALSE, include = FALSE}
library(tidyverse)
library(PerformanceAnalytics)
library(sf)
library(sp)
library(tmap)
library(patchwork)

tidy_asthma = 
  read_csv(file = "./final_data/tidy_asthma.csv")

tidy_asthma 

shape = st_read("./final_data/shapefiles/", quiet = TRUE)

tidy_asthma_sf = merge(shape, tidy_asthma, by.x = "UHFCODE", by.y = "geo_join_id") %>% 
  janitor::clean_names() %>% 
  mutate(uhfcode = as.character(uhfcode),
         time_period = as.character(time_period),
         homes_with_cockroaches = as.numeric(homes_with_cockroaches))

tidy_asthma_sf_ll = st_transform(tidy_asthma_sf, "+proj=longlat +datum=WGS84")
```
```{r, echo = FALSE}

p2 = 
  tidy_asthma_sf_ll %>%
  ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma)) + 
  geom_density() +
  facet_grid(~borough) +
  labs(
    x = "Percent of Public School Children with Asthma",
    y = "Frequency",
    title = "Distribution of Childhood Asthma by Borough")
p2
```

## Cockroach Prevalence by Borough Graph 
```{r, echo = FALSE}
p1 = 
  tidy_asthma_sf_ll %>%
  ggplot(aes(x = homes_with_cockroaches)) + 
  geom_density() +
  facet_grid(~borough) +
  labs(
    x = "Percent of Homes with Cockroaches",
    y = "Frequency",
    title = "Distribution of Cockroaches by Borough"
  ) 

p1
```


### Related work 

## Initial Questions 

## Data 

# Data

* We built our dataset using the [NYC Gov Environmental & Health Data Portal](http://a816-dohbesp.nyc.gov/IndicatorPublic/BuildATable.aspx). The map was built with the UHF code key from [NYC Health](http://a816-dohbesp.nyc.gov/IndicatorPublic/EPHTPDF/uhf42.pdf). The asthma CSV file we used is linked (here)[asthma_data.csv]. We selected all variables relating to asthma, cockroaches, and poverty through string detection. 

* Our github repository is linked [here](https://github.com/prisrinivasan/p8105_final)

## Exploratory analysis

* The graphs and analyses can be found [here](regression.Rmd).
* We started by creating a correlation matrix to assess multicollinearity and baseline correlations
* We then assessed the crude relationship between exposure, cockroach presence in the home and the outcome, childhood asthma
* 

## Discussion

