---
title: "priyanka"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(viridis)
library(sf)
library(sp)
library(rvest)
library(tigris)
library(maptools)
library(ggmap)
library(broom)
library(httr)
library(rgdal)
library(tmap)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Alice's dataset (same code as her Rmd!):
```{r}
all_data = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Homes with Cockroaches", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma"),
        geo_type_name == "UHF42", 
        measure %in% c("Percent", "Rate")
         ) %>% 
  mutate(geo_join_id = as.character(geo_join_id),
         geo_join_id = if_else(geo_place_name == "Bensonhurst - Bay Ridge", "209", geo_join_id),
         geo_join_id = as.factor(geo_join_id)) %>% 
  pivot_wider(
    names_from = name, 
    values_from = data_value
    ) %>%
  mutate(
    time_period = recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011")) %>%
  filter(time_period %in% c("2011", "2012", "2013", "2014")) %>% 
  janitor::clean_names() %>% 
  select(geo_join_id, time_period, measure, indicator_id, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma)

df1 = 
  all_data  %>% 
  select(geo_join_id, time_period, measure, indicator_id, homes_with_cockroaches) %>% 
  drop_na()
df2 = 
all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, poverty) %>% 
  drop_na()
df2
df3 = 
  all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, public_school_children_5_14_yrs_old_with_asthma) %>% 
  drop_na()
df3
df4 = 
  all_data %>% 
  select(geo_join_id, time_period, measure, indicator_id, public_school_children_5_14_yrs_old_with_persistent_asthma) %>% 
  drop_na()
df4
## Joining
tidy_asthma = 
  full_join(df1, df2, by = c("geo_join_id", "time_period")) %>% 
  full_join(df3, by = c("geo_join_id", "time_period")) %>% 
  full_join(df4, by = c("geo_join_id", "time_period")) %>% 
  select(geo_join_id, time_period, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma) %>% 
  view()
  

```

# Shape files
Shape files turned out better. Used a different file from nyc.gov because the Env & Health Data Portal file was inconsistent - combined certain neighborhoods that weren't combined in the rest of the datasets.

[Source of shape files](https://www1.nyc.gov/site/doh/data/health-tools/maps-gis-data-files-for-download.page)

```{r, message=FALSE, warning=FALSE}
shape = st_read("./final_data/shapefiles/", quiet = TRUE)

#look at what we have
#str(shape)

#plot all polygon boundaries
#plot(shape)

#plot just geometry column
plot(st_geometry(shape))
```

Checking the `shape` tibble, it looks like the values we need are stored properly. Merging with the `tidy_asthma`
dataset below.

```{r, message=FALSE, warning=FALSE}
merged <- merge(shape, tidy_asthma, by.x = "UHFCODE", by.y = "geo_join_id")
```


Below is a plot filtered for 2011 poverty values. 

```{r sick plot, message=FALSE, warning=FALSE}
asthma_2011 = merged %>%
  filter(time_period == "2011")

plot(asthma_2011["poverty"])
```

Same with ggplot, which marks latitudes and longitudes:

```{r}
ggplot(asthma_2011) + 
  geom_sf(aes(fill = poverty))
```

Tmap with overlay on real map:

```{r}
tmap_mode("view")

tm_shape(asthma_2011) +
  tm_polygons("homes_with_cockroaches", 
              style = "quantile",
              title = "Poverty Levels in NYC Neighborhoods")
```


