---
title: "priyanka"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(viridis)
library(sf)
library(sp)
library(rvest)
library(tigris)
library(maptools)
library(leaflet)
library(ggmap)
library(broom)
library(httr)
library(rgdal)
library(tmap)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Edits below from Alice's code to:

  * recode `Bensonhurst - Bay Ridge` as `209`
  * add hospitalization variable
  * add geo_place_name variable

These are saved as `tidy_asthma.csv` in our data folder.

```{r tidying data, message=FALSE, include=FALSE}
all_data = 
read.csv(file = "./final_data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Homes with Cockroaches", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma", 
                     "Asthma Hospitalizations (Children 5 to 17 Yrs Old)"),
        geo_type_name %in% c("UHF42", "Borough"), 
        measure %in% c("Percent", "Rate")
         ) %>% 
  mutate(geo_join_id = as.character(geo_join_id),
         geo_join_id = if_else(geo_place_name == "Bensonhurst - Bay Ridge", "209", geo_join_id),
         geo_join_id = as.factor(geo_join_id)) %>% 
  pivot_wider(
    names_from = name, 
    values_from = data_value
    ) %>%
  mutate(time_period = dplyr::recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011")) %>%
  filter(time_period %in% c("2011", "2012", "2013", "2014")) %>% 
  janitor::clean_names() %>% 
  dplyr::select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma, asthma_hospitalizations_children_5_to_17_yrs_old)

df1 = all_data  %>% 
  dplyr::select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, homes_with_cockroaches) %>% 
  drop_na()

df2 = all_data %>% 
  dplyr::select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, poverty) %>% 
  drop_na()

df3 = all_data %>% 
  dplyr::select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, public_school_children_5_14_yrs_old_with_asthma) %>% 
  drop_na()

df4 = all_data %>% 
  dplyr::select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, public_school_children_5_14_yrs_old_with_persistent_asthma) %>% 
  drop_na()

df5 = all_data %>% 
  dplyr::select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  drop_na()

## Joining
tidy_asthma = 
  full_join(df1, df2, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df3, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df4, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df5, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  dplyr::select(geo_type_name, geo_join_id, time_period, geo_place_name, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma, asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  view()
  
#save as csv
write.csv(tidy_asthma,"./final_data/tidy_asthma.csv", row.names = FALSE)
```

# Shape files
I used a different file from nyc.gov because the Env & Health Data Portal file was inconsistent - combined certain neighborhoods that weren't combined in the rest of the datasets.

[Source of shape files](https://www1.nyc.gov/site/doh/data/health-tools/maps-gis-data-files-for-download.page)

```{r, message=FALSE, warning=FALSE}
shape = st_read("./final_data/shapefiles/", quiet = TRUE)

#look at what we have
#str(shape)

#plot all polygon boundaries
#plot(shape)

#plot just geometry column
plot(st_geometry(shape))

tidy_asthma_sf = merge(shape, tidy_asthma, by.x = "UHFCODE", by.y = "geo_join_id") %>% 
  janitor::clean_names() %>% 
  mutate(uhfcode = as.character(uhfcode),
         time_period = as.character(time_period),
         homes_with_cockroaches = as.numeric(homes_with_cockroaches))

#changing geometry to latitude and longitude
tidy_asthma_sf_ll = st_transform(tidy_asthma_sf, "+proj=longlat +datum=WGS84")

#creating color palette Blues for cockroaches
pal = colorQuantile("Blues", tidy_asthma_sf_ll$homes_with_cockroaches, n = 7)
```

Checking the `shape` tibble, it looks like the values we need are stored properly. Merging with the `tidy_asthma`
dataset below.

```{r, message=FALSE, warning=FALSE}
merged <- merge(shape, tidy_asthma, by.x = "UHFCODE", by.y = "geo_join_id")
```


**Leaflet:**
The code chunk is modified into the [shiny app](https://prisrinivasan.shinyapps.io/roach_shiny/).

```{r}
tidy_asthma_sf_ll %>%
    mutate(roach_label = str_c("Homes with Cockroaches: ", homes_with_cockroaches, "\n, Neighborhood: ", uhf_neigh)) %>% 
    filter(time_period == 2011) %>% 
    leaflet() %>%
    addProviderTiles("CartoDB.Positron") %>%
    addPolygons(stroke = TRUE,
                weight = 2,
                opacity = 1,
                color = ~pal(homes_with_cockroaches),
                fillOpacity = 0.8,
                smoothFactor = 0.2,
                highlight = highlightOptions(weight = 3,
                                             color = "white",
                                             bringToFront = TRUE),
                label = ~roach_label) %>% 
    addLegend("bottomright",
              pal = pal,
              values = ~homes_with_cockroaches,
              title = "Percent of Homes with Cockroaches",
              opacity = 1,
              labFormat = function(type, cuts, p) {
                n = length(cuts)
                paste0(round(cuts, 1)[-n], " &ndash; ", round(cuts, 1)[-1], "%")
                })

```


Below is a plot filtered for 2011 poverty values. 

```{r sick plot, message=FALSE, warning=FALSE}
asthma_2011 = merged %>%
  filter(time_period == "2011")

plot(asthma_2011["poverty"])
```

Same with ggplot, which marks latitudes and longitudes:

```{r}
ggplot(asthma_2011) + 
  geom_sf(aes(fill = poverty))
```

Tmap with overlay on real map:

```{r}
tmap_mode("view")

tm_shape(asthma_2011) +
  tm_polygons("homes_with_cockroaches", 
              style = "quantile",
              title = "Poverty Levels in NYC Neighborhoods")
```


The unused ggplot for shiny:
```{r}
#Row

# -----------------------------------------------------------------------

#plot = renderPlot({tidy_asthma_sf_ll %>% 
#    mutate(borough = map(borough, func, input$borough),
#           roach_label = str_c("Homes with Cockroaches: ", #homes_with_cockroaches, "\n , Neighborhood: ", uhf_neigh)) %>% 
#    filter(borough == TRUE,
#           time_period == input$year) %>%
#    ggplot() +
#    geom_sf(aes(fill = homes_with_cockroaches), show.legend = "polygon")})

```


