---
title: "Exploratory Analyses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(PerformanceAnalytics)
```
# Crude Asthma Rate Analysis with Neighborhood, Poverty, and Pest Predictors

Method
The dataset we used was compiled using the Environmental and Health Dataset from the NYC Health Department. We selected variables based on our preliminary hypothesis regarding asthma, poverty, neighborhood, cockroaches, and school children 

# Correlation Matrix
```{r, echo = FALSE, error = TRUE}

chart.Correlation(tidy_asthma[,3:6], histogram = TRUE, method = c("pearson", "kendall", "spearman"))
```
The correlation matrix displays the baseline relationship between each variable. All of the correlations are above 0.77 which indicates strong correlation. As we model build we will be aware of any multicollinearity.

# Cockroach Distribution by Boro
```{r, echo = FALSE}
violin_roaches = 
  tidy_asthma %>%
  filter(time_period == c("2011", "2014")) %>%
  ggplot(aes(x = homes_with_cockroaches, y = public_school_children_5_14_yrs_old_with_asthma)) + 
  geom_violin() +
  facet_grid(~time_period) +
  labs(
    x = "Percent of Homes with Cockroaches",
    y = "Percent of Public School Children with Asthma",
    title = "Percent of School Children with Asthma and Cockroaches in Home per Year"
  )

violin_roaches
```
The cockroach distribution 

# Linear Model of Cockroach Presence, Poverty, and Neighobrhood on Asthma
```{r}

```

