---
title: "Kelly"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(ggridges)
library(patchwork)
library(plotly)
library(ggplot2)
library(gganimate)
library(modelr)
library(mgcv)

```

```{r, error = TRUE}
all_available_data=
  read_csv(file="./final_data/all_available_data.csv")

Asthma_data=
  read_csv(file="./final_data/asthma_data.csv")

mold_data=
  read_csv(file="./final_data/mold_home.csv")
```

### Kelly vision: X : borough name , Y: Cockroach Rate, by different boroughs 
#### Trying to include Borough's in here--using Alice's Method to include Borough name


```{r, error = TRUE}
library(car)
tidy_asthma = 
  read_csv(file = "./final_data/tidy_asthma.csv")

tidy_asthma 
pp = tidy_asthma %>% 
  mutate(
    geo_join_id = as.numeric(geo_join_id) ) %>% 
  ggplot(aes(x = geo_join_id, y = homes_with_cockroaches)) + 
  geom_point(aes(colour = time_period), alpha = 0.5) +
  labs(title = "Prevalence of Cockroaches in Different Neighborhoods",
     x = 'Geo_ID', 
     y = 'Rate of Homes with Cockroaches(%)'
)
#pp

# gganimate parts
#pp + transition_time(time_period) + labs(title="Year:{frame_time}")

#animate(pp, fps = 2, height = 600, width = 1000)
```

## using lorraine's code as a guide
```{r, error = TRUE}
test_prev=
  tidy_asthma %>% 
  drop_na(homes_with_cockroaches) %>% 
  mutate(
    geo_join_id=as.numeric(geo_join_id),
    time_period=as.factor(time_period),
    time_period=fct_relevel(time_period, "2011"="2011", "2012"="2012", "2013" = "2013", "2014" = "2014")
  ) %>% 
  arrange(geo_join_id) %>% 
  plot_ly(
    x=~geo_join_id, 
    y=~homes_with_cockroaches,
    color=~geo_join_id,
    frame=~ time_period,
    hoverinfo= "text",
    type="scatter",
    mode="markers"
  )

test_prev
  
```



```{r, error = TRUE}
## see changes with Hospitalzation by time
set.seed(10)
library(ggplot2)

tidy_asthma=
  read_csv(file="./final_data/tidy_asthma.csv")

ggplot_data = 
tidy_asthma %>%
  filter(time_period == c("2011", "2014"))

plot_tidy = 
ggplot(ggplot_data, aes(x = homes_with_cockroaches, y = asthma_hospitalizations_children_5_to_17_yrs_old, color = homes_with_cockroaches)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  facet_grid(~time_period) +
  labs(x = "Percent of Homes with Cockroaches",
       y = "Percent of School Children Hospitalized from Asthma",
       title = "Crude Association Homes with Cockroaches and Childhood Asthma Hospitalization") +
  viridis::scale_color_viridis(aes(color = homes_with_cockroaches), discrete = FALSE)

plot_tidy

test_prev=
  tidy_asthma %>% 
  drop_na(asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  mutate(
    geo_join_id=as.character(geo_join_id),
   time_period = as.numeric(levels(time_period)) [time_period]
  ) %>% 
  arrange(geo_join_id) %>% 
  plot_ly(
    x=~geo_join_id, 
    y=~asthma_hospitalizations_children_5_to_17_yrs_old,
    color=~geo_join_id,
    frame=~ time_period,
    hoverinfo= "text",
    type="scatter",
    mode="markers"
  )

test_prev
```
<<<<<<< HEAD


## regression model:
```{r, error = TRUE}
#set seed for reproducibility
set.seed(10)
#fitting the full model


tidy_asthma = 
  read_csv(file = "./final_data/tidy_asthma.csv")
test=
  tidy_asthma %>% 
  filter(time_period == c("2011", "2014"))

## Model 0 -- crude 
#model0_fit = lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches, data=test) %>% broom::tidy()




## Model 1
#model1_fit = lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches + poverty + geo_join_id + time_period, data=test) %>% broom::tidy()

#model1_fit

## Model 2
#model2_fit = lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * geo_join_id*time_period, data=test) %>% broom::tidy()


# Looking at residuals -- FAILED
#residual_0=
  #model0_fit %>% 
  #modelr::add_residuals(test, model0_fit) %>% 
  #ggplot(aes(x=public_school_children_5_14_yrs_old_with_asthma, y=resid)) +geom_point()

#residual_1=
  #modelr::add_residuals(test, model1_fit) %>% 
  #ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma, y = resid)) + geom_point()

#residual_2=
  #modelr::add_residuals(tidy_asthma, model2_fit) %>% 
  #ggplot(aes(x = public_school_children_5_14_yrs_old_with_asthma, y = resid)) + geom_point()


```

## Cross Validation
```{r, error = TRUE}
cv_df=
  crossv_mc(tidy_asthma, 100) %>% 
  mutate(
    train=map(train, as_tibble),
    test = map(test, as_tibble)
  )

#cv_df=
 # cv_df %>% 
  #mutate(fit = map(train, ~lm(public_school_children_5_14_yrs_old_with_asthma ~ ., data=.x)),
    #    model1_lm= map(train, ~lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches + poverty + geo_join_id, data=.x)),
     #   model2_lm=map(train, ~lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * geo_join_id, data=.x)),
      #  model5_lm=map(train, ~lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * geo_join_id * asthma_hospitalizations_children_5_to_17_yrs_old, data=.x))) %>% 
  # mutate(rmse_fit = map2_dbl(fit, test, ~rmse(model=.x, data=.y)),
    #     rmse_model1= map2_dbl(model1_lm, test, ~rmse(model=.x, data=.y)),
     #    rmse_model2= map2_dbl(model2_lm, test, ~rmse(model=.x, data=.y)),
      #   rmse_model5=map2_dbl(model5_lm, test, ~rmse(model=.x, data=.y)))
```

## Trying to do multi-linear modeling 
```{r}
library(car)
tidy_asthma = 
  read_csv(file = "./final_data/tidy_asthma.csv")

#tidy_asthma 

#ggplot_data = 
#tidy_asthma %>%
  #filter(time_period == c("2011", "2014"))

#linear_model = 
 # lm(public_school_children_5_14_yrs_old_with_asthma ~ homes_with_cockroaches * poverty * time_period * geo_join_id, data = ggplot_data) %>%
 # broom::glance()
#layout(matrix(c(1,2,3,4),2,2))
#plot(linear_model)
```

